---
title: "BMI 881 literature seminar HW1"
author: "Tuo Wang"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  echo = FALSE,          # don't show code
  collapse = TRUE,
    warning = FALSE,       # don't show warnings
  message = FALSE,       # don't show messages (less serious warnings)
  cache = FALSE,         # set to TRUE to save results from last compilation
  fig.align = "center",  # center figures
  comment = "#>"
)

require(tidyverse)      
```

The data has three categorical variable.

- `year`: 1990, 2005, 2010
- `age_group`: <75, >=75, all
- `income_group`: high, low and middle, all

## First attempt

To visualize the mortality values, I use `geom_point` for mortality rates and `geom_errorbar` for the confidence intervals. To show year, age and income, I could use `color` and `facet_wrap`. Because there are some points too closed to each other, I use `position_dodge` to let the point position deviating a little bit.

```{r, fig.width=8, fig.height=5}
setwd("/Users/tuowang/Documents/UW-Madison Fall 2019/BMI 881/hw1/")
df <- read.csv("feigin2014_table1_mortality.csv")
ggplot(df,aes(x = age_group, y = mortality_rate, color = income_group)) +
  geom_point(position = position_dodge(0.5)) +
  geom_errorbar(aes(ymin = interval_low, ymax=interval_high),size = 1.0,width=.5,position = position_dodge(0.5))+
  facet_wrap(~year, scales = "free_y")+
  scale_color_manual(values = c("#386cb0","#fdb462","#7fc97f"))+
  labs(title = 'Figure 1')+
  theme_bw()
```


```{r, fig.width=8, fig.height=5}
ggplot(df,aes(x = income_group, y = mortality_rate, color = age_group)) +
  geom_point(position = position_dodge(0.1)) +
  geom_errorbar(aes(ymin = interval_low, ymax=interval_high),size = 1.0,width=.5,position = position_dodge(0.1))+
  facet_wrap(~year, scales = "free_y")+
  scale_color_manual(values = c("#386cb0","#fdb462","#7fc97f"))+
  labs(title = 'Figure 2')+
  theme_bw()
```

```{r, fig.width=8, fig.height=5}
ggplot(df, aes(x = year, y = mortality_rate, color = age_group))+
  geom_point()+
  geom_line(size = 1.0,  alpha=0.7)+
  geom_errorbar(aes(ymin = interval_low, ymax=interval_high),size = 1.0,width=1)+
  facet_grid(~income_group, scales = "free_y")+
  scale_color_manual(values = c("#386cb0","#fdb462","#7fc97f"))+
  labs(title = 'Figure 3')+
  theme_bw()
```

## Second attempt

After plotting the two figures above, I realized that the mortality rates in age group '>=75' is much larger than the values in the other group. The confidence interval for the other two age groups are not obvious in `figure1`, `figure2` and `figure3`. So if I want to include all the information, I need a figure that can handle different scale. I created a new categorical variable, called `factor1`. `factor1` equals to 0 if the age group is '>=75' and 1 otherwise. Therefore, `factor1` can distinguish between age group '>=75' and the other two groups. In `figure4`, the left Y-axis is mortality rate, the right Y-axis is `factor1`. I eliminated the grey boxes in the right to get a better looking. The top X-axis is income group and the bottom X-axis is year. You can see the obvious difference between `figure3` and `figure4`.


```{r, fig.width=8, fig.height=6.5}
df$factor1 <- rep(1, 27)
df$factor1[df$age_group == '>=75'] <- 0
ggplot(df, aes(x = year, y = mortality_rate, color = age_group))+
  geom_point()+
  geom_line(size = 1.0,  alpha=0.7)+
  geom_errorbar(aes(ymin = interval_low, ymax=interval_high),size = 1.0,width=1)+
  facet_grid(factor1~income_group, scales = "free_y")+
  scale_color_manual(values = c("#386cb0","#fdb462","#7fc97f"))+
  theme_bw()+
  labs(title = 'Figure 4')+
  theme(strip.text.y = element_blank()) 
```



## Code

```{r ref.label = knitr::all_labels(), echo = TRUE, eval = FALSE}
# this R markdown chunk generates a code appendix
```
